Placing the spotted T Tauri star LkCa 4 on an HR diagram

### 2016-07-27 document/ms.tex
This commit was posted to Overleaf.com as ms_25jul_mgs.tex, along with some updated tables and updated ms.bib.
Prose added by GJH on the Overleaf on July 27, several new figures for his section.

\section{Monitoring the Spots of LkCa 4}

@@

In \S 4.1 we describe how the two components of the photosphere fit low resolution spectra and optical photometry.  The spectral features and colors should change as the spot rotates around the star, so that the visible fraction of hot and cool spot changes.  While the variability in optical brightness is well documented and motivatet this study, in \S 4.2-4.3 we convert the $V$ band brightness to a spot filling factor and then quantify spectroscopic and color variability caused by changes in the spot filling factor.

\begin{figure*}
    	%\vspace{-60mm}
	\centering
	\includegraphics[width=0.95\textwidth]{figures/lkca4_lores} 
	%\vspace{-100mm}
	\caption{The low-resolution optical/near-IR spectrum of LkCa 4 is }
	\label{fig:lores}
\end{figure*}

@@

Figure~\ref{fig:lores} shows that the two-component photosphere matches flux-calibrated spectra well from $\sim 6000-2.5$ $\mu$m.  The two components have fixed temperatures of 4200 K covering 30\% of the visible surface and 3000 K covering 70\% of the visible surface area.  The flux and extinction are then scaled to match the spectrum.

The value for the extinction depends on XXX...

@@

Since 1992 the $V$ band brightness has varied from $12.3$ to $13.2$ mag, with a typical range of $0.5$ mag in any single epoch.  In a two-temperature atmosphere, this brightness corresponds directly to a filling factor for both components.

We model the rotational modulation of LkCa 4 in different photometric bands by fixing the two temperatures to 4200 and 3000 K and setting the 70\% covering fraction of the cool component to $V_{est}=12.85$, as estimated at the time of the IGRINS spectrum.  The colors for the two components are obtained from the main sequence colors compiled by \citet{kenyon95}.  The colors from dwarfs stars should have minimal contribution from spots; measurements of pre-main sequence colors will include some unknown contribution of spots and cannot be used for this analysis.

In these simple models, the blue emission is dominated by the hotter component, while both components contribute to the red and infrared emission.   
The $0.5$ mag range in $V$-band brightness requires at least a factor of 1.6 change in the visible surface area of hot component.  Figure~\ref{fig:spotcolors} shows the

@@

\subsection{Phased colors}

\begin{figure*}
	\centering
	%\vspace{-60mm}
	\includegraphics[width=0.95\textwidth]{figures/lkca4_tio_forpaper} 
    %	\vspace{-30mm}
	\caption{Variability in TiO bands measured with ESPaDOnS (see also Table X).  The V-band emission is estimated from fits to the ASAS-SN lightcurve obtained during the same period.  The main panel shows a correlation between V-band magnitude and the TiO-7140 index, while the inset shows a similar correlation with the average of the TiO 6200, CaH 6800, and TiO 7600 indices.}
	\label{fig:tiovar}
\end{figure*}

@@

\begin{figure*}
	\centering
    %	\vspace{-30mm}
	\includegraphics[width=0.95\textwidth]{figures/lkca4_hrdiag}
    %	\vspace{-30mm}
	\caption{Placeholder for HR Diagram Figure}
	\label{fig:hrdiag}
\end{figure*}

@@

Table REF lists measurements of LkCa4 from previous studies.
\input{tables/tbl_history.tex}

>>>>

\section{Spots at Low Spectral Resolution}

@@

In \S 4.1 we describe how the two components of the photosphere fit low resolution spectra and optical/near-IR photometry.  The brightness, colors, spectral features should change as the spot rotates around the star.  In \S 4.2, we convert the $V$ brightness to a filling factor of the cool spot, and then in \S 4.2-4.3 we compare how the expected rotational modulation of photometry and spectral indices compares with observed measurements.

\begin{figure*}
	\centering
    	\vspace{-120mm}
	\includegraphics[width=0.95\textwidth]{figures/lkca4_lores} 
	\caption{The low-resolution optical/near-IR spectrum of LkCa 4 is }
	\label{fig:lores}
\end{figure*}

\begin{figure*}
	\centering
    	\vspace{-120mm}
	\includegraphics[width=0.95\textwidth]{figures/lkca4_lores_twopanel} 
	\caption{The low-resolution optical/near-IR spectrum of LkCa 4 is }
	\label{fig:lores_twopanel}
\end{figure*}

\begin{figure*}
	\centering
    	\vspace{-20mm}
	\includegraphics[width=0.95\textwidth]{figures/vband_spot} 
    	\vspace{-100mm}
\caption{The low-resolution optical/near-IR spectrum of LkCa 4 is }
\label{fig:vband_spot}
\end{figure*}

@@

Figure~\ref{fig:lores} shows that the two-component photosphere matches flux-calibrated spectra well from $\sim 0.6-2.5$ $\mu$m.  The two components have fixed temperatures of 4200 K covering 30\% of the visible surface and 3000 K covering 70\% of the visible surface area.  The flux and extinction are then scaled to match the spectrum.

The lower panels show photospheric spectra scaled separately to the observed red-optical and near-IR spectra.  The single temperature spectra are unable to reproduce the observed emission.  The $3900$ K component fits the TiO bands at reasonably well at $<7400$ \AA\ but fails to reproduce molecular bands at longer wavelengths.  The $3500$ K component hotter component suffers from the opposite problem, yielding molecular bands at short wavelengths that are too deep.  On the best-fit photosphere from \S 3 fits both the optical and the near-IR bands well.

SED...

The value for the extinction depends on.

Relative to main sequence colors, the $E(B-V)=0.24$ would yield.

%Furlan:  K7, V-IC=1.92, E(V-I)=1.6 => 0.85 mag;  close enough to their 1.1 mag
%V-K = 4.18  ; MS colors:  3.16 (KH95) => Av=1.1 mag

%V-J:  2.8-3.8
%J-K:  0.9-1.0

 The upper limit may be estimated by comparing the observed $B-V$ color to main sequence $B-V$ colors.

The $B-V$

The $J-K=0.93$ color is equal to the expected color to within 0.05 => 0.3 mag (for T=3000, 4200)

\begin{figure*}
	\centering
	\includegraphics[width=0.49\textwidth]{figures/bv_kh95} 
    \includegraphics[width=0.49\textwidth]{figures/vr_kh95} 
\caption{The low-resolution optical/near-IR spectrum of LkCa 4 is }
\vspace{-50mm}
\label{fig:colors}
\end{figure*}

@@

Since 1992 the $V$ band brightness has varied from $12.3$ to $13.2$ mag, with a typical range of $0.5$ mag in any single epoch.  In a two-temperature atmosphere with fixed temperatures, this brightness corresponds directly to a filling factor for both components.  Figure~\ref{vband_spot} shows this conversion for temperatures of 4200 K, 3000 K, and a 75\% filling factor for the 3000 K component at $V_{est}=12.83$, as estimated at the time of the IGRINS spectrum.

The model photometry is obtained from the main sequence colors compiled by \citet{kenyon95}.  Colors from dwarfs stars should have minimal contribution from spots; measurements of pre-main sequence colors will include some unknown contribution of spots and cannot be used for this analysis.  Synthetic spectra of $3000-4000$ K stars predict $B-V$ colors that are much bluer than observed and that become bluer at cooler temperatures, the opposite of what is observed.  As a consequence, synthetic photometry also cannot be used here.

The optical emission is dominated by the hotter component while both components contribute to the red and infrared emission (see Figure \ref{fig:lores}).  The $0.7$ mag range in $V$-band brightness requires at a factor of 1.9 change in the visible surface area of hot component in the limit that the cool component does not contribute any $V$-band emission.  In the 2014-2015 ASAS-SN photometry, the minimum and maximum brightness correspond to cool spot coverage of 83\% and 68\%, respectively.  LkCa 4 was brighter in past epochs, which corresponds to either a larger covering fraction for the hot component (as assumed here) or components with hotter temperatures.

The rotational modulation leads to much smaller brightness changes at near-IR wavelengths than at optical wavelengths (see Table~\ref{tab:photrange}).  The smaller amplitude of near-IR brightness is also seen in a few spotted stars in the optical/IR monitoring of NGC 2264 \citep{cody14}, although those stars have a much smaller amplitude than LkCa 4 in $V$ band.  Figure~\ref{fig:colors} demonstrates that the $B-V$ color from \citet{grankin08} does not depend on $V$, while the $V-R_J$ becomes slightly redder when $V$ is smaller.  The predicted colors from rotational modulation do not match the observed $V-R_J$ colors or the correlation with $V$, either because of changes in spot properties since the \citet{grankin08} data were obtained or because a two-temperature model for the photosphere is overly simplistic.

\begin{table}
\caption{Predicted Photometric Range - based on $V$, adapt for $A_V$}
\label{tab:photrange}
\begin{tabular}{cccc}
\hline
Band & Low Spot & High Spot & Range\\
Cool Spot & 68\% & 83\% &\\
\hline
$B$ & 13.80 & 14.48 & 0.68\\
$V$ & 12.56 & 13.23 & 0.67\\
$R$ & 11.53 & 12.17 & 0.64 \\
$I$ & 10.80 & 11.38 & 0.58\\
$J$ & 9.79 & 10.00 & 0.21\\
$K$ & 8.89 & 9.04 & 0.15\\
\hline
\end{tabular}
\end{table}

@@

\begin{figure*}
	\centering
	\vspace{-60mm}
	\includegraphics[width=0.95\textwidth]{figures/lkca4_tio_forpaper} 
    	\vspace{-30mm}
	\caption{Variability in TiO bands measured with ESPaDOnS (see also Table X).  The V-band emission is estimated from fits to the ASAS-SN lightcurve obtained during the same period.  The main panel shows a correlation between V-band magnitude and the TiO-7140 index, while the inset shows a similar correlation with the average of the TiO 6200, CaH 6800, and TiO 7600 indices.}
	\label{fig:tiovar}
\end{figure*}

@@

\begin{figure*}
	\centering
    %	\vspace{-30mm}
	\includegraphics[width=0.95\textwidth]{figures/lkca4_hrdiag_forpaper}
    	\vspace{-80mm}
	\caption{HR Diagram Figure}
	\label{fig:hrdiag}
\end{figure*}

@@

Table \ref{tbl_history} lists measurements of LkCa4 from previous studies.
\input{tables/tbl_history.tex}
